---
title: "Patterns of Religious/Spiritual Engagement Among Sexual and Gender Minorities of Color"
short-title: "R/S Engagement Among SGMs of Color"
author: "Seungju Kim"
institute: "University of Illinois"
short-institute: "UIUC"
department: "Department of Psychology"
acknowledgment: "This material is based upon work supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. DGE 21-46756. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation."
format:
  beamer:
    template: template.tex
    outertheme: miniframes
    aspectratio: 169
    section-titles: true
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
pacman::p_load(here, tidyverse)
```

# Background {.unlisted .unnumbered}

## Health Disparities in SGMs 

## Religiousness/Spirituality and SGM Health

# Present Study {.unlisted .unnumbered}

## Participants and Procedure
Hello 

## Measures

## Analysis Plan

# Results {.unlisted .unnumbered}

## Latent Profile Analysis

```{r echo=FALSE}
#| fig-width: 20
#| fig-height: 9
#| out-width: "200%"
#| out-height: "85%"
#| fig-align: "center"
lpa_plot <- readRDS(file = here("analysis", "figures", "lpa_plot.Rds"))

line_data <- ggplot_build(lpa_plot)$data[[2]]

line_styles <- line_data %>%
  group_by(group) %>%
  summarize(
    color = first(colour),
    linetype = first(linetype)
  )

p <- lpa_plot
for (i in seq_len(nrow(line_styles))) {
  g <- line_styles$group[i]
  p <- p + geom_line(
    data = filter(line_data, group == g),
    aes(x = x, y = y, group = group),
    color = line_styles$color[i],
    linetype = line_styles$linetype[i],
    linewidth = 5,
    show.legend = FALSE,
    inherit.aes = FALSE
  )
}

p +
  theme(
    # Make all text larger
    text = element_text(size = 18),           # Base text size
    axis.title = element_text(size = 30),     # Axis titles
    axis.text = element_text(size = 26),      # Axis labels
    legend.title = element_text(size = 28),   # Legend title
    legend.text = element_text(size = 26),    # Legend text
    plot.title = element_text(size = 22)) +    # Plot title (if you have one)
  guides(
    colour = guide_legend(override.aes = list(linewidth = 5, size = 4)),
    linetype = guide_legend(override.aes = list(linewidth = 5, size = 4)),
    shape = guide_legend(override.aes = list(linewidth = 5, size = 4))
  )
```


## Demographic Predictors

## Outcomes Pt. 1

## Outcomes Pt. 2

# Implications {.unlisted .unnumbered}

## R/S matter to SGMs

## R/S engagement is linked to health and identity development

## Future directions