---
title: "LPA with POC Sample"
format:
  html: 
    toc: true
    toc-depth: 6
    toc-location: left
    embed-resources: true
editor: visual
code-fold: true
code-summary: "Show the code"
editor_options: 
  chunk_output_type: console
execute:
  warning: false
---

# Set Up

## Library and Processing Settings

```{r, output=FALSE, message = FALSE, warning = FALSE}
# install.packages("pacman") # install if you don't have this package
pacman::p_load(tidySEM, haven, tidyverse, future, progressr, gt, nnet, OpenMx, gtsummary, mice, here, emmeans)
set.seed(152738)
options(mc.cores = 8)
plan(multisession) 
handlers("progress") 
```

## Import Data

```{r}
org_data <- read_sav(here("data", "prri_final_shareable.sav"))
data <- org_data
```

## Prepare Data

```{r}
names(data) <- tolower(names(data))

data$poc <- rowSums(data[, c("black", "americanindian", "asianamerican", "latine", "mena")])

data$racsum <- rowSums(data[, c("white", "black", "americanindian", "asianamerican", "latine", "mena")])

# Subset the data, where if POC then keep, otherwise remove.
data <- data %>%
  filter(poc > 0)

# ref_data <- data

data <- data %>%
  filter(!(responseid %in% c("R_BzwG6ehjWl0zgRz", "R_1nPF8RGMaq0IFME")))
# These cases were removed due to some errors with the LPA model running.

data <- rename(data,
               "filter" = "filter_$")
# data <- data[-141, ] 

data <- data %>%
  mutate(racem = case_when(
    racsum > 1 ~ "Multi-Racial",
    black == 1 ~ "Black/African American",
    americanindian == 1 ~ "American Indian",
    asianamerican == 1 ~ "Asian/Asian American",
    latine == 1 ~ "Hispanic/Latin",
    mena == 1 ~ "Middle Easter/North African",
    other == 1 ~ "Another Identity Not Listed",
    TRUE ~ NA_character_
  ))

data$degreeofconflict <- ifelse(data$conflict == 0, 0, data$degreeofconflict)
table(data$degreeofconflict)

data$sex <- ifelse(data$sex == 1, 0, 1)


data$religious <- ifelse(data$religiousaffiliation %in% c(1:8, 12, 13, 14), 2, 1)
# First, identify which columns are character types
char_cols <- sapply(data, is.character)

# Then, only replace empty strings with NA in character columns
data[char_cols][data[char_cols] == ""] <- NA

# First, convert haven_labelled variables to regular R types
data_unlabelled <- data

# Convert all haven_labelled columns to appropriate R types
for(col in names(data_unlabelled)) {
  if(inherits(data_unlabelled[[col]], "haven_labelled")) {
    data_unlabelled[[col]] <- as.vector(data_unlabelled[[col]])
  }
}
data_subset <- data_unlabelled[, !names(data_unlabelled) %in% c("tnbextra", "state", "urbanicity", "income")]

data_imputed <- mice::complete(mice::mice(data_subset, 
                                         m = 20))
```

## Preprocess for LPA

```{r}
lpa_data <- subset(data, select = c("relcom", 
                                    "relconnect",
                                    "serviceattendance", 
                                    "nora"))



# lpa_data <- lpa_data[-141, ]
# data <- data[-141, ]

# This case caused issues and would not allow the model to run for some reason EVEN though TidySEM handles missing data with FIML.
```

## Descriptives

```{r}
descriptives(lpa_data) %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt() %>%
  tab_header(
    title = "Descriptives"
  ) 
```

## Run LPA

```{r}
set.seed(123)
# lpa1 <- mx_profiles(data = lpa_data, classes = 1:7)
# saveRDS(lpa1, here("analysis", "model_objects", "lpa1_0424a_poc.RData"))
lpa1 <- readRDS(here("analysis", "model_objects", "lpa1_0424a_poc.RData"))
```

::: callout-note
Missing data was automatically handled using the Full-Information Maximum Likelihood (FIML) estimator.
:::

::: callout-warning
We have previously run the analyses and will only present them from previous loads. You are welcome to run them, however they are computationally taxing and can take more than an hour.
:::

## Fit Indices

```{r}
fit <- table_fit(lpa1)
fit %>%
  select(Name, LL, Parameters, AIC, BIC, Entropy,
         prob_min, prob_max, n_min, np_ratio, np_local) %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt() %>%
  tab_header(
    title = "Profiles Fit Indices"
  ) 

plot(fit)
```

## LRT

```{r}
# lrt_lpa <- lr_lmr(lpa1)
# saveRDS(lrt_lpa, here("0225_Analysis", "lrt_lpa.RData"))
# lrt_lpa <- readRDS(here("0225_Analysis", "lrt_lpa.RData"))
# lrt_lpa
```

## BLRT

```{r}
# lpa1_blrt <- BLRT(lpa1, replications = 1000)
# saveRDS(lpa1_blrt, here("0225_Analysis", "lpa1_blrt.RData"))
# lpa1_blrt <- readRDS(here("0225_Analysis", "lpa1_blrt.RData")) # then load it again
# lpa1_blrt
```

## Plots of Profiles

```{r, include=FALSE}
lpa1_plot <- plot_profiles(
  lpa1,
  ci = NULL,
  sd = FALSE,
  add_line = TRUE,
  rawdata = FALSE,
  bw = FALSE
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
lpa1_plot
```

::: callout-note
In contrast to the previous models, this model usese FIML which allows for high quality results (i.e. greater Entropy across the board). Evidence points to many things, so this one leans more upon rationale. Pretty much each class solution has good separation (Entropy \>.90).

Results suggest Class 5 is the best fit.
:::

## LPA 5-Class Solution Data

```{r}
lpa1_solution <- lpa1[[4]]
lpa1_class4 <- class_prob(lpa1[[4]])
predicted_class <- lpa1_class4$individual[, ncol(lpa1_class4$individual)]
lpa1_data <- cbind(data, predicted_class)

lpa1_data <- rename(lpa1_data,
                    "class" = "predicted_class") # NEW = OLD
```

## Plot

```{r, include=FALSE}
lpa1_plot <- plot_profiles(
  lpa1_solution,
  ci = NULL,
  sd = FALSE,
  add_line = TRUE,
  rawdata = FALSE,
  bw = F
)

new_labels <- c(
"Rel. Commitment",
"Rel. Community",
"Service Attendance",
"Spiritual Activity"
)

lpa1_plot <- lpa1_plot +
scale_x_continuous(
breaks = 1:4, # Assuming you have 5 variables
labels = new_labels
)
```

```{r}
# Now rename the legend labels with the labels you want
lpa1_plot <- lpa1_plot +
  scale_color_discrete(name = "Profile", 
                      labels = c("Not Religious/Not Spiritual", 
                                 "Religious/Not Spiritual", 
                                 "Not Religius/Spiritual", 
                                 "Religious/Spiritual")) +
  scale_shape_discrete(name = "Profile", 
                      labels = c("Not Religious/Not Spiritual", 
                                 "Religious/Not Spiritual", 
                                 "Not Religius/Spiritual", 
                                 "Religious/Spiritual")) +
  scale_linetype_discrete(name = "Profile", 
                         labels = c("Not Religious/Not Spiritual", 
                                 "Religious/Not Spiritual", 
                                 "Not Religius/Spiritual", 
                                 "Religious/Spiritual"))

lpa1_plot <- lpa1_plot +
  theme(
    text = element_text(size = 14),                     # Base font size for all text
    axis.title = element_text(size = 16),               # Axis titles
    axis.text = element_text(size = 12),                # Axis text
    legend.title = element_text(size = 16),             # Legend title
    legend.text = element_text(size = 12),              # Legend text
    plot.title = element_text(size = 18)                # Plot title if you have one
  )

# Display the plot
lpa1_plot
```

## Demographics

```{r}
selected_vars <- c("sex", "gender", "age", "education", "racem", "income", "sexualidentity", "urbanicity", "class")
dem_data <- subset(lpa1_data, select = selected_vars)

# Ensure all variables are properly formatted
dem_data <- dem_data %>%
  mutate(across(c(sex, gender, education, income, racem, sexualidentity, urbanicity, class), as_factor),
         age = as.numeric(age))

# Create the summary table
dem_table <- gtsummary::tbl_summary(
  dem_data,
  by = class,
  include = c(age, sex, gender, racem, sexualidentity, education, income, urbanicity),
  missing = "no",
  type = list(age ~ "continuous2"),
  statistic = list(age ~ "{mean} ({sd})"),
  digits = everything() ~ 2,
  label = list(
    age ~ "Age",
    sex ~ "Sex Assigned at Birth",
    gender ~ "Gender Identity",
    racem ~ "Racial Identity",
    sexualidentity ~ "Sexual Orientation",
    income ~ "Financial Hardship",
    education ~ "Education",
    urbanicity ~ "Residence"
  )
)

dem_table

# dem_table %>%
#   as_gt() %>%
#   gt::gtsave(filename = here("0225_Analysis", "lpa_dem_table.docx"))
```

## Predictors
```{r}
calculate_pairwise_ors <- function(pseudo_class_obj, predictor_name) {
  # Extract results from the pseudo-class object
  results_df <- pseudo_class_obj
  
  # Filter for the specified predictor
  predictor_rows <- results_df[results_df$term == predictor_name, ]
  
  # Extract classes, coefficients and standard errors
  classes <- as.character(predictor_rows$y.level)
  coefs <- predictor_rows$estimate
  se <- predictor_rows$std.error
  p_values <- predictor_rows$p.value
  
  # Create a data frame with odds ratios for each class vs reference
  or_vs_ref <- data.frame(
    Class = classes,
    Reference = "Reference Class",
    Coefficient = coefs,
    OR = exp(coefs),
    OR_CI_Lower = exp(coefs - 1.96 * se),
    OR_CI_Upper = exp(coefs + 1.96 * se),
    p_value = p_values
  )
  
  # Calculate pairwise comparisons between all classes
  n_classes <- length(classes)
  pairwise_results <- data.frame(
    Class1 = character(),
    Class2 = character(),
    Coefficient_Diff = numeric(),
    SE = numeric(),
    z_value = numeric(),
    p_value = numeric(),
    OR = numeric(),
    OR_CI_Lower = numeric(),
    OR_CI_Upper = numeric(),
    stringsAsFactors = FALSE
  )
  
  if (n_classes > 1) {
    for (i in 1:(n_classes-1)) {
      for (j in (i+1):n_classes) {
        # Calculate difference in coefficients
        coef_diff <- coefs[i] - coefs[j]
        
        # Calculate SE for the difference (using error propagation)
        diff_se <- sqrt(se[i]^2 + se[j]^2)
        
        # Calculate z-value and p-value
        z_val <- coef_diff / diff_se
        p_val <- 2 * pnorm(-abs(z_val))
        
        # Calculate OR and confidence intervals
        or <- exp(coef_diff)
        or_lower <- exp(coef_diff - 1.96 * diff_se)
        or_upper <- exp(coef_diff + 1.96 * diff_se)
        
        # Add to results data frame
        pairwise_results <- rbind(pairwise_results, data.frame(
          Class1 = classes[i],
          Class2 = classes[j],
          Coefficient_Diff = coef_diff,
          SE = diff_se,
          z_value = z_val,
          p_value = p_val,
          OR = or,
          OR_CI_Lower = or_lower,
          OR_CI_Upper = or_upper
        ))
      }
    }
  }
  
  # Return both sets of results
  return(list(
    vs_reference = or_vs_ref,
    pairwise = pairwise_results
  ))
}

```


### Age

```{r, output=FALSE, message = FALSE, warning = FALSE}
age_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ age, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)

ors_age <- calculate_pairwise_ors(age_log, "age")

# Print results
ors_age$vs_reference

```

```{r}
age_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Residence

```{r, output=FALSE, message = FALSE, warning = FALSE}
city_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ urbanicity, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
city_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Sex

```{r, output=FALSE, message = FALSE, warning = FALSE}
sex_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ sex, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
sex_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Race

#### Asian

```{r, output=FALSE, message = FALSE, warning = FALSE}
asian_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ asianamerican, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
asian_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

#### Latine

```{r, output=FALSE, message = FALSE, warning = FALSE}
latine_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ latine, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
latine_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

#### Black

```{r, output=FALSE, message = FALSE, warning = FALSE}
black_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ black, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
black_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

#### Multiracial

```{r, output=FALSE, message = FALSE, warning = FALSE}
mult_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ multiracial, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
mult_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Gender Identity

```{r, output=FALSE, message = FALSE, warning = FALSE}
tnb_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ tnb, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
tnb_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Sexual Orientation

```{r, output=FALSE, message = FALSE, warning = FALSE}
plur_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ plurisexual, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
plur_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Identity Conflict
```{r}
lpa1_data$degreeofconflict <- as.numeric(lpa1_data$degreeofconflict)
doc_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ degreeofconflict, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)

ors_doc <- calculate_pairwise_ors(doc_log, "degreeofconflict")

# Print results
ors_doc$vs_reference
```


### Interpersonal Struggles
```{r}
irs_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ irs, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)

irs_doc <- calculate_pairwise_ors(irs_log, "irs")

# Print results
irs_doc$vs_reference
```


## BCH/Outcomes

The recommended method of handling distal outcomes with LPA.

### Demographics

```{r}
selected_vars <- c("dep", "ih", "lgbtconnect", "relconnect", "degreeofconflict", "irs", "class")
dem2_data <- subset(lpa1_data, select = selected_vars)


# Create the summary table
dem2_table <- gtsummary::tbl_summary(
  dem2_data,
  by = class,
  include = c(dep, ih, lgbtconnect, relconnect, degreeofconflict, irs),
  missing = "no",
  type = list(
    dep ~ "continuous2",
    ih ~ "continuous2",
    lgbtconnect ~ "continuous2",
    relconnect ~ "continuous2",
    degreeofconflict ~ "continuous2",
    irs ~ "continuous2"
  ),
  statistic = list(
    dep ~ "{mean} ({sd})",
    ih ~ "{mean} ({sd})",
    lgbtconnect ~ "{mean} ({sd})",
    relconnect ~ "{mean} ({sd})",
    degreeofconflict ~ "{mean} ({sd})",
    irs ~ "{mean} ({sd})"
  ),
  digits = list(
    dep ~ 2,
    ih ~ 2,
    lgbtconnect ~ 2,
    relconnect ~ 2,
    degreeofconflict ~ 2,
    irs ~ 2
  ),
  label = list(
    dep ~ "Depression",
    ih ~ "Internalized Homonegativity",
    lgbtconnect ~ "LGBT Connect",
    relconnect ~ "Religious Connect",
    degreeofconflict ~ "Degree of Conflict",
    irs ~ "IRS"
  )
)

dem2_table
```

### Depression

```{r}
BCH(lpa1_solution, data = lpa1_data$dep) %>%
  lr_test()
```

### Internalized Homonegativity

```{r}
BCH(lpa1_solution, data = lpa1_data$ih) %>%
  lr_test()
```

### LGBT Connectedness

```{r}
BCH(lpa1_solution, data = lpa1_data$lgbtconnect) %>%
  lr_test()
```

### Rel Connectedness

```{r}
BCH(lpa1_solution, data = lpa1_data$relconnect) %>%
  lr_test()
```

### Degree of Conflict

```{r}
lpa1_data$degreeofconflict <- as.numeric(lpa1_data$degreeofconflict)
BCH(lpa1_solution, data = lpa1_data$degreeofconflict) %>%
  lr_test()
```

### Interpersonal Religious Struggles

```{r}
BCH(lpa1_solution, data = lpa1_data$irs) %>%
  lr_test()
```

```{r}
# covariates_model <- BCH(lpa1_solution,
#                         model = "education",
#                         data = lpa1_data)
# check <- mxCheckIdentification(covariates_model, details=TRUE)
# # Examine results
# summary(covariates_model)
# 
# # Likelihood ratio test for overall effect of covariates
# lr_test(covariates_model)
# 
# # Get odds ratios for covariates
# exp(coef(covariates_model))
```
