---
title: "LPA with POC Sample: LGBT Identity and DUREL"
format:
  html: 
    toc: true
    toc-depth: 6
    toc-location: left
    embed-resources: true
editor: visual
code-fold: true
code-summary: "Show the code"
editor_options: 
  chunk_output_type: console
execute:
  warning: false
---

# Set Up

## Library and Processing Settings

```{r, output=FALSE, message = FALSE, warning = FALSE}
# install.packages("pacman") # install if you don't have this package
pacman::p_load(tidySEM, haven, tidyverse, future, progressr, gt, nnet, OpenMx, gtsummary, mice, here)
set.seed(152738)
options(mc.cores = 8)
plan(multisession) 
handlers("progress") 
```

## Import Data

```{r}
org_data <- read_sav(here("data", "prri_final_shareable.sav"))
data <- org_data
```

## Prepare Data

```{r}
names(data) <- tolower(names(data))

data$poc <- rowSums(data[, c("black", "americanindian", "asianamerican", "latine", "mena")])

data$racsum <- rowSums(data[, c("white", "black", "americanindian", "asianamerican", "latine", "mena")])

# Subset the data, where if POC then keep, otherwise remove.
data <- data %>%
  filter(poc > 0)

# ref_data <- data

data <- data %>%
  filter(!(responseid %in% c("R_BzwG6ehjWl0zgRz", "R_1nPF8RGMaq0IFME")))
# These cases were removed due to some errors with the LPA model running.

data <- rename(data,
               "filter" = "filter_$")
# data <- data[-141, ] 

data <- data %>%
  mutate(racem = case_when(
    racsum > 1 ~ "Multi-Racial",
    black == 1 ~ "Black/African American",
    americanindian == 1 ~ "American Indian",
    asianamerican == 1 ~ "Asian/Asian American",
    latine == 1 ~ "Hispanic/Latin",
    mena == 1 ~ "Middle Easter/North African",
    other == 1 ~ "Another Identity Not Listed",
    TRUE ~ NA_character_
  ))

data$degreeofconflict <- ifelse(data$conflict == 0, 0, data$degreeofconflict)
table(data$degreeofconflict)

data$sex <- ifelse(data$sex == 1, 0, 1)
```

## Preprocess for LPA

```{r}
lpa_data <- subset(data, select = c("sexualidentitysalience",
                                    "relcom", 
                                    "serviceattendance", 
                                    "nora"))

lpa_data <- lpa_data %>%
  mutate(across(where(is.labelled), as.numeric))

lpa_data[lpa_data == ""] <- NA
lpa_data <- lpa_data[rowSums(!is.na(lpa_data)) > 0, ]


lpa_data <- lpa_data[-141, ]
data <- data[-141, ]
# This case caused issues and would not allow the model to run for some reason EVEN though TidySEM handles missing data with FIML.
```

## Descriptives

```{r}
descriptives(lpa_data) %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt() %>%
  tab_header(
    title = "Descriptives"
  ) 
```

## Run LPA

```{r}
# lpa1 <- mx_profiles(data = lpa_data, classes = 1:5)
# saveRDS(lpa1, here("analysis", "model_objects", "lpa_model.RData"))
lpa1 <- readRDS(here("analysis", "model_objects", "lpa_model.RData"))
```

::: callout-note
Missing data was automatically handled using the Full-Information Maximum Likelihood (FIML) estimator.
:::

::: callout-warning
We have previously run the analyses and will only present them from previous loads. You are welcome to run them, however they are computationally taxing and can take more than an hour.
:::

## Fit Indices

```{r}
fit <- table_fit(lpa1)
fit %>%
  select(Name, LL, Parameters, AIC, BIC, Entropy,
         prob_min, prob_max, n_min, np_ratio, np_local) %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt() %>%
  tab_header(
    title = "Profiles Fit Indices"
  ) 

plot(fit)
```

## LRT

```{r}
# lrt_lpa <- lr_lmr(lpa1)
# saveRDS(lrt_lpa, here("identity_and_durel", "lrt_lpa.RData"))
# lrt_lpa <- readRDS(here("identity_and_durel", "lrt_lpa.RData"))
# lrt_lpa
```

## BLRT

```{r}
# lpa1_blrt <- BLRT(lpa1, replications = 1000)
# saveRDS(lpa1_blrt, here("identity_and_durel", "lpa1_blrt.RData"))
# lpa1_blrt <- readRDS(here("identity_and_durel", "lpa1_blrt.RData")) # then load it again
# lpa1_blrt
```

## Plots of Profiles

```{r, include=FALSE}
lpa1_plot <- plot_profiles(
  lpa1,
  ci = NULL,
  sd = FALSE,
  add_line = TRUE,
  rawdata = FALSE,
  bw = FALSE
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
lpa1_plot
```

::: callout-note
In contrast to the previous models, this model usese FIML which allows for high quality results (i.e. greater Entropy across the board). Evidence points to many things, so this one leans more upon rationale. Pretty much each class solution has good separation (Entropy \>.90).

Results suggest Class 4 is the best fit.
:::

## LPA 4-Class Solution Data

```{r}
lpa1_solution <- lpa1[[4]]
lpa1_class4 <- class_prob(lpa1[[4]])
predicted_class <- lpa1_class4$individual[, ncol(lpa1_class4$individual)]
lpa1_data <- cbind(data, predicted_class)

lpa1_data <- rename(lpa1_data,
                    "class" = "predicted_class") # NEW = OLD
```

## Plot

```{r, include=FALSE}
lpa1_plot <- plot_profiles(
  lpa1_solution,
  ci = NULL,
  sd = FALSE,
  add_line = TRUE,
  rawdata = FALSE,
  bw = F
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
lpa1_plot
```

## Demographics

```{r}
selected_vars <- c("sex", "gender", "age", "education", "racem", "income", "sexualidentity", "urbanicity", "class")
dem_data <- subset(lpa1_data, select = selected_vars)

# Ensure all variables are properly formatted
dem_data <- dem_data %>%
  mutate(across(c(sex, gender, education, income, racem, sexualidentity, urbanicity, class), as_factor),
         age = as.numeric(age))

# Create the summary table
dem_table <- gtsummary::tbl_summary(
  dem_data,
  by = class,
  include = c(age, sex, gender, racem, sexualidentity, education, income, urbanicity),
  missing = "no",
  type = list(age ~ "continuous2"),
  statistic = list(age ~ "{mean} ({sd})"),
  digits = everything() ~ 2,
  label = list(
    age ~ "Age",
    sex ~ "Sex Assigned at Birth",
    gender ~ "Gender Identity",
    racem ~ "Racial Identity",
    sexualidentity ~ "Sexual Orientation",
    income ~ "Financial Hardship",
    education ~ "Education",
    urbanicity ~ "Residence"
  )
)

dem_table

# dem_table %>%
#   as_gt() %>%
#   gt::gtsave(filename = here("identity_and_durel", "lpa_dem_table.docx"))
```

## Predictors

### Age

```{r, output=FALSE, message = FALSE, warning = FALSE}
age_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ age, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
age_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Residence

```{r, output=FALSE, message = FALSE, warning = FALSE}
city_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ urbanicity, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
city_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Sex

```{r, output=FALSE, message = FALSE, warning = FALSE}
sex_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ sex, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
sex_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Race

#### Asian

```{r, output=FALSE, message = FALSE, warning = FALSE}
asian_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ asianamerican, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
asian_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

#### Latine

```{r, output=FALSE, message = FALSE, warning = FALSE}
latine_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ latine, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
latine_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

#### Black

```{r, output=FALSE, message = FALSE, warning = FALSE}
black_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ black, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
black_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

#### Multiracial

```{r, output=FALSE, message = FALSE, warning = FALSE}
mult_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    
    m <- nnet::multinom(class ~ multiracial, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
mult_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Gender Identity

```{r, output=FALSE, message = FALSE, warning = FALSE}
tnb_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ tnb, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
tnb_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

### Sexual Orientation

```{r, output=FALSE, message = FALSE, warning = FALSE}
plur_log <- tidySEM::pseudo_class(
  x = lpa1_solution, 
  model = function(data) {
    m <- nnet::multinom(class ~ plurisexual, data = data)
    class(m) <- head(class(m), 1)
    m
  }, 
  data = lpa1_data  # Pass your actual data here
)
```

```{r}
plur_log %>%
  mutate(across(where(is.numeric), ~ round(., digits = 3))) %>%
  gt::gt()
```

## BCH/Outcomes

The recommended method of handling distal outcomes with LPA.

### Demographics

```{r}
selected_vars <- c("dep", "ih", "lgbtconnect", "relconnect", "degreeofconflict", "irs", "class")
dem2_data <- subset(lpa1_data, select = selected_vars)


# Create the summary table
dem2_table <- gtsummary::tbl_summary(
  dem2_data,
  by = class,
  include = c(dep, ih, lgbtconnect, relconnect, degreeofconflict, irs),
  missing = "no",
  type = list(
    dep ~ "continuous2",
    ih ~ "continuous2",
    lgbtconnect ~ "continuous2",
    relconnect ~ "continuous2",
    degreeofconflict ~ "continuous2",
    irs ~ "continuous2"
  ),
  statistic = list(
    dep ~ "{mean} ({sd})",
    ih ~ "{mean} ({sd})",
    lgbtconnect ~ "{mean} ({sd})",
    relconnect ~ "{mean} ({sd})",
    degreeofconflict ~ "{mean} ({sd})",
    irs ~ "{mean} ({sd})"
  ),
  digits = list(
    dep ~ 2,
    ih ~ 2,
    lgbtconnect ~ 2,
    relconnect ~ 2,
    degreeofconflict ~ 2,
    irs ~ 2
  ),
  label = list(
    dep ~ "Depression",
    ih ~ "Internalized Homonegativity",
    lgbtconnect ~ "LGBT Connect",
    relconnect ~ "Religious Connect",
    degreeofconflict ~ "Degree of Conflict",
    irs ~ "IRS"
  )
)

dem2_table
```

### Depression

```{r}
BCH(lpa1_solution, data = lpa1_data$dep) %>%
  lr_test()
```

### Internalized Homonegativity

```{r}
BCH(lpa1_solution, data = lpa1_data$ih) %>%
  lr_test()
```

### LGBT Connectedness

```{r}
BCH(lpa1_solution, data = lpa1_data$lgbtconnect) %>%
  lr_test()
```

### Rel Connectedness

```{r}
BCH(lpa1_solution, data = lpa1_data$relconnect) %>%
  lr_test()
```

### Degree of Conflict

```{r}
lpa1_data$degreeofconflict <- as.numeric(lpa1_data$degreeofconflict)
BCH(lpa1_solution, data = lpa1_data$degreeofconflict) %>%
  lr_test()
```

### Interpersonal Religious Struggles

```{r}
BCH(lpa1_solution, data = lpa1_data$irs) %>%
  lr_test()
```

```{r}
# covariates_model <- BCH(lpa1_solution,
#                         model = "education",
#                         data = lpa1_data)
# check <- mxCheckIdentification(covariates_model, details=TRUE)
# # Examine results
# summary(covariates_model)
# 
# # Likelihood ratio test for overall effect of covariates
# lr_test(covariates_model)
# 
# # Get odds ratios for covariates
# exp(coef(covariates_model))
```
